\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{extarrows}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, calc, decorations.pathreplacing, decorations.pathmorphing, fit, backgrounds}

\setlength{\emergencystretch}{1em}

\newcommand{\excursus}[2]{%
  \begin{center}
  \begin{tikzpicture}
    \node[draw=blue!40, fill=blue!4, rounded corners=4pt,
          inner sep=10pt, text width=0.88\textwidth, font=\small]
    {\textbf{#1}\;\;---\;\;#2};
  \end{tikzpicture}
  \end{center}}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue,
  citecolor=blue
}

\title{Real-Time Web Communication Standards:\\A Retrospective and Projection\\2021--2031}
\author{}
\date{March 2026}

\begin{document}
\maketitle

\tableofcontents
\listoffigures
\listoftables
\newpage

\section*{Executive Summary}
\addcontentsline{toc}{section}{Executive Summary}

\begin{itemize}
  \item \textbf{Late 2021:} Only WebRTC was a finalized standard with broad
    browser support, and even it had major cross-browser gaps (Safari: H.264
    only, no simulcast, broken iOS DataChannel). QUIC was newly standardized
    but not universal. HTTP/3 was pre-RFC. WebTransport was non-functional.
    WebCodecs existed in one browser. MoQ did not exist.

  \item \textbf{March 2026:} The core stack (WebRTC, QUIC, HTTP/3) is
    standardized and universal. WebCodecs is cross-browser with 94\% coverage.
    Two gaps remain: Safari lacks WebTransport (82\% coverage) and Encoded
    Transforms for E2EE remain Chrome-only.

  \item \textbf{2028--2029 projection:} WebTransport reaches Safari via
    Interop~2026. Encoded Transforms become cross-browser. WebCodecs achieves
    W3C Recommendation. AV1 becomes a practical WebRTC codec as hardware
    encoding expands. MoQ and RTP over QUIC reach RFC status.

  \item \textbf{2029--2031 projection:} Full stack convergence. The
    ``unbundled'' architecture (WebTransport + WebCodecs + WebAssembly) becomes
    viable for specialized use cases. AV1 hardware encoding is ubiquitous.
    MoQ enables standards-based sub-second streaming with CDN fan-out.

  \item \textbf{Hardware and network:} Standards maturation compounds with
    hardware improvement (60--80\% CPU gains vs.\ 2021, ubiquitous AV1 silicon)
    and network growth (200--300~Mbps fixed, 100+~Mbps mobile). By 2030, a
    720p call consumes less than 0.5\% of link capacity.

  \item \textbf{Information-theoretic view (Appendix~\ref{app:info-theory}):}
    Codec efficiency gaps narrow from H.264 ($\Delta \approx 2.1$) to AV1
    ($\Delta \approx 0.4$), approaching the Shannon rate-distortion bound.
    QUIC eliminates head-of-line blocking, recovering up to 34\% per-stream
    throughput at 5\% loss. WebTransport's unreliable datagrams outperform
    WebSocket for any application with a freshness half-life below 34~ms.
\end{itemize}

\section{Introduction}

This document examines the concrete browser capabilities enabled by the
real-time web communication standards collected in the \texttt{standards/}
directory of this project. It compares what these standards made possible in
browsers as of late 2021 with what they enable as of March 2026, then projects
where the stack is heading through 2031. The discussion focuses exclusively on
what standards adoption in browsers enabled and will enable --- not on market
trends, product adoption, or commercial developments.

The standards under review span five areas:
\begin{itemize}
  \item \textbf{WebRTC} --- W3C WebRTC 1.0 and supporting IETF protocols
    (RFCs 8829, 8445, 8826, 8827, 3550, 3711)
  \item \textbf{QUIC} --- RFC 9000
  \item \textbf{HTTP/3} --- RFC 9114
  \item \textbf{WebTransport} --- W3C Working Draft and RFC 9297
  \item \textbf{WebCodecs} --- W3C Working Draft
\end{itemize}

Together, these standards define a full stack for real-time communication in
the browser: from low-level transport (QUIC) through application-layer
protocols (HTTP/3, WebTransport) to peer-to-peer media and data exchange
(WebRTC) and direct codec access (WebCodecs).

Sections~3--7 detail each standard's browser capabilities in late 2021 and
March 2026. Sections~8--9 project the stack's evolution through 2031, grounded
in active W3C charter milestones, IETF draft timelines, browser vendor
signals, and hardware and network trends. Section~10 summarizes the full
2021--2031 trajectory. Appendix~\ref{app:info-theory} applies
information-theoretic measures to quantify channel capacity, source coding
efficiency, and effective information rates across the same time horizons.

Section~2 first provides a concise overview of what each standard does and how
the pieces fit together, for readers unfamiliar with the protocol landscape.

\section{The Standards Stack}
\label{sec:standards-stack}

The standards under review form a layered architecture built on two
complementary transport paths---one \emph{peer-to-peer} (browser talks
directly to browser) and one \emph{client-server} (browser talks to a
server)---with a shared codec layer that bridges both.  This section explains
what each standard does, what problem it solves, and how the pieces fit
together.  Figure~\ref{fig:protocol-stack} provides a visual map; the
subsections below walk through each layer from the bottom up.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  block/.style={draw, rounded corners=2pt, minimum height=0.65cm,
    align=center, font=\footnotesize, thick},
  transport/.style={block, fill=blue!10},
  security/.style={block, fill=red!6},
  session/.style={block, fill=green!10},
  app/.style={block, fill=orange!10},
  codec/.style={block, fill=purple!8},
  lbl/.style={font=\scriptsize\itshape, text=black!50},
  arr/.style={->, thick, gray!50},
]

% --- WebRTC side (left) ---
\node[lbl] at (-3.5, 5.6) {Peer-to-peer path};

\node[transport, minimum width=4.2cm] (udp1) at (-3.5, 0) {UDP};
\node[session, minimum width=4.2cm]   (ice)  at (-3.5, 1.0) {ICE (RFC 8445)};
\node[security, minimum width=4.2cm]  (dtls) at (-3.5, 2.0) {DTLS-SRTP (RFCs 8827, 3711)};
\node[session, minimum width=4.2cm]   (rtp)  at (-3.5, 3.0) {RTP / RTCP (RFC 3550)};
\node[session, minimum width=4.2cm]   (jsep) at (-3.5, 4.0) {JSEP / SDP (RFC 9429)};
\node[app, minimum width=4.2cm]       (rtc)  at (-3.5, 5.0) {RTCPeerConnection (W3C)};

% --- QUIC side (right) ---
\node[lbl] at (3.5, 5.6) {Client-server path};

\node[transport, minimum width=4.2cm] (udp2) at (3.5, 0) {UDP};
\node[transport, minimum width=4.2cm] (quic) at (3.5, 1.5) {QUIC + TLS\,1.3 (RFC 9000)};
\node[session, minimum width=4.2cm]   (h3)   at (3.5, 3.0) {HTTP/3 (RFC 9114)};
\node[app, minimum width=4.2cm]       (wt)   at (3.5, 4.5) {WebTransport (RFC 9297 + W3C)};

% --- WebCodecs (centre, spanning both) ---
\node[codec, minimum width=3.4cm] (wc) at (0, 6.3) {WebCodecs (W3C)};

% --- Vertical arrows (WebRTC) ---
\draw[arr] (udp1) -- (ice);
\draw[arr] (ice)  -- (dtls);
\draw[arr] (dtls) -- (rtp);
\draw[arr] (rtp)  -- (jsep);
\draw[arr] (jsep) -- (rtc);

% --- Vertical arrows (QUIC) ---
\draw[arr] (udp2) -- (quic);
\draw[arr] (quic) -- (h3);
\draw[arr] (h3)   -- (wt);

% --- WebCodecs connections (dashed) ---
\draw[->, thick, purple!40, dashed]
  ($(wc.south west)!0.15!(wc.south)$) -- ($(rtc.north east)!0.15!(rtc.north)$);
\draw[->, thick, purple!40, dashed]
  ($(wc.south east)!0.15!(wc.south)$) -- ($(wt.north west)!0.15!(wt.north)$);

\node[font=\tiny, text=purple!50, anchor=south east] at (-0.5, 5.65)
  {Encoded Transforms};
\node[font=\tiny, text=purple!50, anchor=south west] at (0.5, 5.55)
  {Custom pipelines};

% --- Layer annotations (far left) ---
\draw[decorate, decoration={brace, amplitude=4pt, mirror},
  thick, black!30] (-6.1, -0.2) -- (-6.1, 0.2)
  node[midway, left=5pt, lbl] {Transport};
\draw[decorate, decoration={brace, amplitude=4pt, mirror},
  thick, black!30] (-6.1, 0.8) -- (-6.1, 3.2)
  node[midway, left=5pt, lbl, align=right] {Security,\\Session};
\draw[decorate, decoration={brace, amplitude=4pt, mirror},
  thick, black!30] (-6.1, 3.8) -- (-6.1, 5.2)
  node[midway, left=5pt, lbl] {Application};

\end{tikzpicture}
\caption{Protocol stack for the real-time web communication standards.
  Left: the WebRTC peer-to-peer path.  Right: the QUIC-based client-server
  path.  WebCodecs (top centre) provides frame-level codec access regardless
  of transport.}
\label{fig:protocol-stack}
\end{figure}

\subsection{The Underlying Transport: UDP}

All the real-time protocols discussed here ultimately send their data over
\textbf{UDP (User Datagram Protocol)}.  UDP is intentionally simple: it
sends individual packets (called \emph{datagrams}) from one machine to
another with no guarantees about ordering, delivery, or retransmission.  If a
packet is lost, UDP does not notice and does not resend it.

This might seem like a poor foundation, but for real-time media it is
actually an advantage.  In a video call, a video frame that arrives 200~ms
late is useless---the moment it was needed has already passed.  The
alternative, \textbf{TCP (Transmission Control Protocol)}, guarantees that
every byte arrives in order by automatically retransmitting lost packets, but
this creates a problem called \emph{head-of-line blocking}: while TCP waits
for a lost packet to be retransmitted, it holds up delivery of all
subsequent data, even data for completely independent streams.  For
interactive media, this stalling is worse than simply dropping the lost
packet and moving on.  UDP lets the protocols built on top of it make their
own choices about what to retransmit and what to discard.

\subsection{WebRTC: Peer-to-Peer Media in the Browser}

\textbf{WebRTC (Web Real-Time Communication)} is a W3C standard that gives
browsers a built-in ability to send audio, video, and arbitrary data directly
from one browser to another---peer-to-peer, without routing media through a
server.  The browser exposes this through a single JavaScript API called
\texttt{RTCPeerConnection}.  Under the hood, \texttt{RTCPeerConnection}
orchestrates a set of IETF protocols that handle four problems:
\emph{finding a network path} between two peers, \emph{agreeing what media
to send}, \emph{encrypting the media}, and \emph{actually carrying it}.
The subsections below explain each problem and the protocol that solves it.

\subsubsection{Finding a Path: ICE, STUN, and TURN}

The first problem is deceptively hard.  Most devices on the internet do not
have their own public IP address.  Instead, they sit behind a
\textbf{NAT (Network Address Translator)}---typically the home router or
corporate firewall---which assigns the device a \emph{private} address
(like \texttt{192.168.1.x}) and translates outgoing traffic onto a single
\emph{public} address that is visible to the rest of the internet.  This
means two browsers behind different NATs cannot simply send packets to each
other: neither knows the other's routable address, and even if it did, the
receiving NAT would likely drop the unsolicited incoming packet.

WebRTC solves this with \textbf{ICE (Interactive Connectivity Establishment,
RFC~8445)}.  ICE is not a server---it is an algorithm that runs in each
browser, systematically searching for a working network path between two
peers.  The search uses two types of infrastructure servers:

\begin{description}
  \item[STUN (Session Traversal Utilities for NAT)] is a lightweight,
    stateless server that serves one purpose: when a browser sends it a
    request, the STUN server replies with the browser's \emph{public} IP
    address and port---the address as seen from the far side of the NAT.
    This is called a \emph{server-reflexive candidate}: the browser now knows
    what address the outside world sees for it.  The STUN server relays no
    media and maintains no per-session state; it simply reflects the
    address back.

  \item[TURN (Traversal Using Relays around NAT)] is a heavier-weight server
    used as a last resort.  When two peers are both behind restrictive NATs
    (such as corporate firewalls that block all unsolicited inbound traffic),
    no direct path exists between them.  In this case, both peers connect to a
    TURN server, which allocates ports, authenticates each peer, and
    \emph{relays} all media packets between them for the duration of the
    session.  This works through any NAT because both peers are making
    outbound connections to the TURN server.  The cost is that all media flows
    through the relay, adding latency and bandwidth consumption.
    Approximately 10--20\% of WebRTC connections require TURN.
\end{description}

The ICE process works as follows.  Each browser gathers a set of
\emph{candidates}---potential addresses through which it might be
reachable:

\begin{enumerate}
  \item \textbf{Host candidates:} the device's own local IP addresses
    (useful when both peers are on the same local network).
  \item \textbf{Server-reflexive candidates:} the public address discovered
    via STUN (useful when the NAT allows direct traffic once the mapping is
    known).
  \item \textbf{Relay candidates:} a port allocated on a TURN server
    (the guaranteed fallback).
\end{enumerate}

Both peers exchange their candidate lists (via the signaling mechanism
described below), and ICE constructs every possible pairing of one local
candidate with one remote candidate.  It then tests each pair in priority
order---host pairs first (lowest latency), then server-reflexive pairs, then
relay pairs---by sending a probe and waiting for a response.  The first pair
that works is selected, and media begins flowing over it.

\textbf{Trickle ICE (RFC~8838)} improves this process by allowing candidates
to be exchanged and tested incrementally, as they are discovered, rather
than waiting for the full list to be compiled.  In practice, this reduces
connection setup time to roughly 500--800~ms because the fastest candidates
(host and server-reflexive) are typically discovered and tested within the
first few hundred milliseconds.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  peer/.style={rectangle, draw, rounded corners=2pt, fill=orange!10,
    minimum width=1.6cm, minimum height=0.65cm, font=\small, thick},
  nat/.style={rectangle, draw=red!50, fill=red!6,
    minimum width=0.9cm, minimum height=1.8cm,
    font=\scriptsize, align=center},
  server/.style={rectangle, draw, rounded corners=2pt, fill=#1,
    minimum width=2cm, minimum height=0.65cm, font=\small, thick},
  server/.default={blue!8},
  arr/.style={->, thick, #1},
  arr/.default={black!60},
  darr/.style={->, thick, dashed, #1},
  darr/.default={black!60},
  lbl/.style={font=\scriptsize, fill=white, inner sep=1pt},
]

% --- Left: The NAT problem ---
\node[font=\small\bfseries, anchor=south] at (2.5, 3.6) {The problem};

\node[peer] (pA) at (0, 1.5) {Browser A};
\node[nat]  (nA) at (2, 1.5) {NAT};
\node[nat]  (nB) at (6.5, 1.5) {NAT};
\node[peer] (pB) at (8.5, 1.5) {Browser B};

\node[font=\scriptsize, text=gray] at (0, 0.7) {192.168.1.5};
\node[font=\scriptsize, text=gray] at (8.5, 0.7) {10.0.0.12};

\draw[thick, red!50, decorate,
  decoration={zigzag, segment length=3mm, amplitude=1.2mm}]
  (nA.east) -- (nB.west);
\node[font=\normalsize\bfseries, text=red!60] at (4.25, 1.55) {$\times$};
\node[font=\scriptsize, text=red!60] at (4.25, 2.2) {No direct path};
\node[font=\scriptsize, text=gray] at (4.25, 0.7) {Public internet};

% --- Right: The solution ---
\node[font=\small\bfseries, anchor=south] at (2.5, -0.6) {The solution};

\node[peer] (pA2) at (0, -2) {Browser A};
\node[peer] (pB2) at (8.5, -2) {Browser B};
\node[server=blue!8] (stun) at (4.25, -0.8) {STUN};
\node[server=green!10] (turn) at (4.25, -3.5) {TURN relay};

% STUN arrows
\draw[arr=blue!50] (pA2.north) -- node[lbl, above, sloped, pos=0.35]
  {\scriptsize discover public IP} (stun.south west);
\draw[arr=blue!50] (pB2.north) -- node[lbl, above, sloped, pos=0.35]
  {\scriptsize discover public IP} (stun.south east);

% Direct path (80-90%)
\draw[arr=green!50!black, line width=1.2pt] (pA2.east) --
  node[lbl, above] {\scriptsize direct path (80--90\%)} (pB2.west);

% TURN fallback (10-20%)
\draw[darr=orange!70!black] (pA2.south) |-
  node[lbl, below, pos=0.25] {\scriptsize relay fallback} (turn.west);
\draw[darr=orange!70!black] (turn.east) -|
  node[lbl, below, pos=0.75] {\scriptsize (10--20\%)} (pB2.south);

% Legend
\node[font=\scriptsize, text=gray, align=left, anchor=north west]
  at (-0.5, -4.3) {%
  \tikz[baseline=-0.5ex]{\draw[->, thick, green!50!black, line width=1.2pt]
    (0,0) -- (0.6,0);}~preferred (STUN-assisted) \qquad
  \tikz[baseline=-0.5ex]{\draw[->, thick, dashed, orange!70!black]
    (0,0) -- (0.6,0);}~fallback (TURN relay)};

\end{tikzpicture}
\caption{The NAT problem and how ICE solves it.  Top: two browsers behind
  NATs cannot reach each other directly---neither knows the other's public
  address.  Bottom: STUN discovers each browser's public address; ICE
  attempts a direct connection (successful 80--90\% of the time) and falls
  back to a TURN relay when direct connectivity is impossible.}
\label{fig:nat-ice}
\end{figure}

\subsubsection{Agreeing What to Send: SDP, JSEP, and Signaling}

Before media can flow, two peers must agree on what they will send each
other: which codecs to use (H.264, VP9, AV1), which media types (audio,
video, screen-share), at what resolutions and bitrates, and which encryption
parameters.  This negotiation uses the \textbf{Session Description Protocol
(SDP)}---a text-based format that describes a peer's media capabilities.

The process works through an \emph{offer/answer} exchange:
\begin{enumerate}
  \item \textbf{Peer~A creates an SDP offer} listing everything it can send
    and receive: ``I support H.264 and VP9 video, Opus audio, and here is my
    encryption certificate fingerprint.''
  \item \textbf{The offer is delivered to Peer~B} through a signaling
    channel.
  \item \textbf{Peer~B creates an SDP answer} selecting the common ground:
    ``I also support VP9 and Opus; let's use those.  Here is my encryption
    fingerprint.''
  \item \textbf{The answer is delivered back to Peer~A.}  Both peers now know
    exactly what media they will exchange and how it will be encrypted.
\end{enumerate}

\textbf{JSEP (JavaScript Session Establishment Protocol, RFC~9429;
originally RFC~8829)} defines how the browser creates and processes these
SDP messages.  It specifies the JavaScript API calls---\texttt{createOffer()},
\texttt{createAnswer()}, \texttt{setLocalDescription()},
\texttt{setRemoteDescription()}---that applications use to drive the
negotiation.

Critically, JSEP \emph{deliberately does not specify how} the SDP offer and
answer are delivered between the two peers.  This delivery mechanism is
called \textbf{signaling}, and it is the one part of a WebRTC system that
the application must build itself.  The signaling channel can be anything:
a WebSocket connection, an HTTP request, a message passed through a shared
database---whatever gets the SDP text from one peer to the other.  The
signaling server also relays ICE candidates as they are discovered (for
Trickle ICE).  Importantly, the signaling server never sees or touches the
actual media; it handles only lightweight control messages.

\subsubsection{Encrypting the Media: DTLS and SRTP}

All WebRTC media is encrypted.  There is no option to send unencrypted
audio or video---encryption is mandatory by design.

The encryption uses two protocols working together:

\begin{description}
  \item[DTLS (Datagram Transport Layer Security, RFC~8827)] is essentially
    TLS (the same encryption used for HTTPS websites) adapted for UDP.  Once
    ICE has found a working path between two peers, DTLS performs a
    cryptographic handshake over that path.  During the handshake, each peer
    presents a self-signed certificate, and the remote peer verifies that the
    certificate matches the fingerprint it received earlier in the SDP
    exchange.  This binding between the signaling channel and the media
    channel prevents \emph{man-in-the-middle attacks}: an attacker cannot
    substitute their own encryption keys without the mismatch being detected.
    The DTLS handshake produces a shared set of encryption keys.

  \item[SRTP (Secure Real-time Transport Protocol, RFC~3711)] uses the keys
    negotiated by DTLS to encrypt every individual media packet.  SRTP adds
    a small authentication tag to each packet (a few bytes), enabling the
    receiver to verify that the packet has not been tampered with in transit.
    The overhead is negligible and introduces no meaningful latency.
\end{description}

\subsubsection{Carrying the Media: RTP and RTCP}

With a path established (ICE), capabilities agreed (SDP/JSEP), and
encryption keys negotiated (DTLS), media can finally flow.  The protocol
that carries the actual audio and video is \textbf{RTP (Real-time Transport
Protocol, RFC~3550)}, sent as encrypted SRTP packets.

RTP provides the essential bookkeeping for real-time media:
\begin{itemize}
  \item \textbf{Sequence numbers} so the receiver can detect lost packets
    and reorder packets that arrive out of sequence.
  \item \textbf{Timestamps} so the receiver can synchronize audio and video
    playback (lip sync) and play back frames at the correct rate.
  \item \textbf{Payload type identifiers} so the receiver knows which codec
    was used to encode each packet.
\end{itemize}

Alongside RTP, the \textbf{RTP Control Protocol (RTCP)} provides a feedback
channel.  Receivers periodically send RTCP reports back to the sender with
statistics on packet loss rate, jitter (variation in packet arrival times),
and round-trip time.  The sender uses this feedback to adapt in real time:
if packet loss increases, the video encoder can reduce its bitrate to avoid
further congestion.  This adaptive behaviour is essential for maintaining
call quality over varying network conditions.

\subsubsection{The WebRTC Security Model}

RFC~8826 defines the overarching security principles for WebRTC:
\begin{itemize}
  \item \textbf{Mandatory encryption:} as described above, all media is
    encrypted via DTLS-SRTP.
  \item \textbf{User consent:} the browser must obtain explicit user
    permission before accessing the camera or microphone.  A web page cannot
    silently record audio or video.
  \item \textbf{IP address privacy:} browsers use \textbf{mDNS (multicast
    DNS)} to mask local IP addresses in ICE candidates.  Without this,
    a web page could discover the user's local network address merely by
    initiating a WebRTC connection, creating a privacy leak.  With mDNS,
    the browser substitutes a random hostname that resolves only on the
    local network.
\end{itemize}

\subsubsection{Putting It Together: A WebRTC Call Sequence}

Figure~\ref{fig:webrtc-sequence} illustrates the end-to-end sequence.
When a user joins a video call:

\begin{enumerate}
  \item The browser gathers ICE candidates (local addresses, STUN-discovered
    public addresses, and optionally TURN relay addresses).
  \item The application creates an SDP offer describing the browser's media
    capabilities and sends it, along with the ICE candidates, through the
    signaling server to the remote peer.
  \item The remote peer responds with an SDP answer and its own ICE
    candidates.
  \item ICE tests candidate pairs and selects the best working path.
  \item DTLS performs a handshake over that path to negotiate encryption
    keys.
  \item Encrypted media (audio and video as SRTP packets) begins flowing
    directly between the two browsers.  The signaling server is no longer
    in the media path.
\end{enumerate}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  entity/.style={rectangle, draw, rounded corners=2pt, minimum height=0.7cm,
    align=center, font=\small, thick, fill=#1},
  entity/.default={white},
  msg/.style={->, thick, #1},
  msg/.default={black!70},
  mlbl/.style={font=\scriptsize, fill=white, inner sep=1.5pt},
  timeline/.style={thick, gray!40, dashed},
]

% --- Entities ---
\node[entity=orange!10] (peerA) at (0, 0)    {Browser A};
\node[entity=green!10]  (sig)   at (5, 0)    {Signaling\\Server};
\node[entity=orange!10] (peerB) at (10, 0)   {Browser B};
\node[entity=blue!8]    (stun)  at (12.5, 0) {STUN};

% --- Timelines ---
\draw[timeline] (0, -0.5) -- (0, -10.8);
\draw[timeline] (5, -0.5) -- (5, -10.8);
\draw[timeline] (10, -0.5) -- (10, -10.8);
\draw[timeline] (12.5, -0.5) -- (12.5, -4.5);

% --- Step 1: STUN discovery ---
\draw[msg] (0, -1.2) -- node[mlbl, above, sloped]
  {1.\ STUN request} (12.5, -1.6);
\draw[msg=blue!60] (12.5, -1.8) -- node[mlbl, above, sloped]
  {public IP reply} (0, -2.2);
\draw[msg] (10, -1.4) -- node[mlbl, above, sloped]
  {1.\ STUN request} (12.5, -1.6);
\draw[msg=blue!60] (12.5, -2.0) -- node[mlbl, above, sloped]
  {public IP reply} (10, -2.4);

\node[font=\scriptsize\itshape, text=black!50, anchor=west] at (-3.2, -1.7)
  {Gather candidates};

% --- Step 2: SDP offer + candidates ---
\draw[msg=green!50!black] (0, -3.2) -- node[mlbl, above]
  {2.\ SDP offer + ICE candidates} (5, -3.5);
\draw[msg=green!50!black] (5, -3.7) -- node[mlbl, above]
  {forward to B} (10, -4.0);

\node[font=\scriptsize\itshape, text=black!50, anchor=west] at (-3.2, -3.5)
  {Signaling};

% --- Step 3: SDP answer + candidates ---
\draw[msg=green!50!black] (10, -4.5) -- node[mlbl, above]
  {3.\ SDP answer + ICE candidates} (5, -4.8);
\draw[msg=green!50!black] (5, -5.0) -- node[mlbl, above]
  {forward to A} (0, -5.3);

% --- Step 4: ICE connectivity checks ---
\draw[msg=orange!70!black, <->] (0, -6.2) -- node[mlbl, above]
  {4.\ ICE connectivity checks (probe/response)} (10, -6.2);

\node[font=\scriptsize\itshape, text=black!50, anchor=west] at (-3.2, -6.2)
  {Path selection};

% --- Step 5: DTLS handshake ---
\draw[msg=red!50!black, <->] (0, -7.4) -- node[mlbl, above]
  {5.\ DTLS handshake (key exchange)} (10, -7.4);

\node[font=\scriptsize\itshape, text=black!50, anchor=west] at (-3.2, -7.4)
  {Encryption};

% --- Step 6: Media flow ---
\draw[msg=blue!60, <->, line width=1.2pt] (0, -8.8) -- node[mlbl, above]
  {6.\ Encrypted media (SRTP audio + video)} (10, -8.8);
\draw[msg=blue!40, <->, dashed] (0, -9.4) -- node[mlbl, above]
  {RTCP feedback (loss, jitter, RTT)} (10, -9.4);

\node[font=\scriptsize\itshape, text=black!50, anchor=west] at (-3.2, -9.1)
  {Media flow};

% --- Signaling server not in media path ---
\node[font=\scriptsize\itshape, text=gray, align=center] at (5, -9.5)
  {Signaling server\\not in media path};

% Brace for direct peer-to-peer
\draw[decorate, decoration={brace, amplitude=5pt},
  thick, blue!40] (10.4, -5.8) -- (10.4, -9.8)
  node[midway, right=6pt, font=\scriptsize, text=blue!60, align=left]
  {Direct\\peer-to-peer};

\end{tikzpicture}
\caption{WebRTC call establishment sequence.  Steps~1--3 use supporting
  infrastructure (STUN server, signaling server).  Once ICE selects a path
  and DTLS establishes encryption (steps~4--5), media flows directly between
  the two browsers (step~6).  The signaling server plays no further role
  in the media path.}
\label{fig:webrtc-sequence}
\end{figure}

\subsection{QUIC: A Modern Transport Protocol}

\textbf{QUIC (RFC~9000)} is a general-purpose transport protocol that, like
WebRTC's media path, runs over UDP.  It was originally developed by Google
and standardized by the IETF in 2021.  QUIC is designed to solve several
long-standing problems with TCP, the transport protocol that has carried most
internet traffic for decades.

\textbf{The head-of-line blocking problem.}  TCP provides a single,
ordered byte stream: every byte must be delivered in sequence.  When a
browser opens multiple parallel data flows over one TCP connection (as
HTTP/2 does), a single lost packet blocks \emph{all} of those flows until
the lost packet is retransmitted and received.  This is head-of-line
blocking.  QUIC solves it by providing \emph{many independent streams}
within a single connection, each with its own sequencing and flow control.
A lost packet on one stream stalls only that stream; all others continue
unaffected.  For applications carrying multiple concurrent flows---such as
a video call with several media tracks---this independence is critical.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  pkt/.style={minimum width=0.55cm, minimum height=0.45cm,
    font=\tiny, align=center, thick, draw},
  lost/.style={pkt, fill=red!25, draw=red!60},
  ok/.style={pkt, fill=#1, draw=gray!60},
  ok/.default={blue!15},
  blocked/.style={pkt, fill=gray!15, draw=gray!40, densely dotted},
  lbl/.style={font=\scriptsize},
  heading/.style={font=\small\bfseries},
]

% === TCP side (left) ===
\node[heading] at (2.5, 3.6) {TCP: single ordered stream};

% Stream labels
\node[lbl, anchor=east] at (-0.15, 2.2) {Video};
\node[lbl, anchor=east] at (-0.15, 1.4) {Audio};
\node[lbl, anchor=east] at (-0.15, 0.6) {Data};

% Timeline arrow
\draw[->, thick, gray!40] (0, -0.1) -- (5.5, -0.1)
  node[anchor=north, font=\tiny, text=gray] {time};

% Video packets
\node[ok=blue!15] at (0.5, 2.2) {};
\node[ok=blue!15] at (1.2, 2.2) {};
\node[lost] at (1.9, 2.2) {\bfseries $\times$};
\node[blocked] at (2.6, 2.2) {};
\node[blocked] at (3.3, 2.2) {};
\node[blocked] at (4.0, 2.2) {};
\node[blocked] at (4.7, 2.2) {};

% Audio packets
\node[ok=green!15] at (0.5, 1.4) {};
\node[ok=green!15] at (1.2, 1.4) {};
\node[blocked] at (1.9, 1.4) {};
\node[blocked] at (2.6, 1.4) {};
\node[blocked] at (3.3, 1.4) {};
\node[blocked] at (4.0, 1.4) {};
\node[blocked] at (4.7, 1.4) {};

% Data packets
\node[ok=orange!15] at (0.5, 0.6) {};
\node[ok=orange!15] at (1.2, 0.6) {};
\node[blocked] at (1.9, 0.6) {};
\node[blocked] at (2.6, 0.6) {};
\node[blocked] at (3.3, 0.6) {};
\node[blocked] at (4.0, 0.6) {};
\node[blocked] at (4.7, 0.6) {};

% Blocked annotation
\draw[thick, red!60, decorate,
  decoration={brace, amplitude=4pt}]
  (1.65, 2.65) -- (5.0, 2.65)
  node[midway, above=4pt, font=\scriptsize, text=red!60]
  {all streams blocked};

% Retransmit annotation
\draw[->, thick, red!50, dashed] (1.9, 1.9) -- (1.9, -0.1)
  node[anchor=north, font=\tiny, text=red!50] {retransmit wait};

% === QUIC side (right) ===
\node[heading] at (9.5, 3.6) {QUIC: independent streams};

\node[lbl, anchor=east] at (6.85, 2.2) {Video};
\node[lbl, anchor=east] at (6.85, 1.4) {Audio};
\node[lbl, anchor=east] at (6.85, 0.6) {Data};

% Timeline arrow
\draw[->, thick, gray!40] (7, -0.1) -- (12.5, -0.1)
  node[anchor=north, font=\tiny, text=gray] {time};

% Video packets - one lost, only video stalls
\node[ok=blue!15] at (7.5, 2.2) {};
\node[ok=blue!15] at (8.2, 2.2) {};
\node[lost] at (8.9, 2.2) {\bfseries $\times$};
\node[blocked] at (9.6, 2.2) {};
\node[blocked] at (10.3, 2.2) {};
\node[ok=blue!15] at (11.0, 2.2) {};
\node[ok=blue!15] at (11.7, 2.2) {};

% Audio packets - unaffected
\node[ok=green!15] at (7.5, 1.4) {};
\node[ok=green!15] at (8.2, 1.4) {};
\node[ok=green!15] at (8.9, 1.4) {};
\node[ok=green!15] at (9.6, 1.4) {};
\node[ok=green!15] at (10.3, 1.4) {};
\node[ok=green!15] at (11.0, 1.4) {};
\node[ok=green!15] at (11.7, 1.4) {};

% Data packets - unaffected
\node[ok=orange!15] at (7.5, 0.6) {};
\node[ok=orange!15] at (8.2, 0.6) {};
\node[ok=orange!15] at (8.9, 0.6) {};
\node[ok=orange!15] at (9.6, 0.6) {};
\node[ok=orange!15] at (10.3, 0.6) {};
\node[ok=orange!15] at (11.0, 0.6) {};
\node[ok=orange!15] at (11.7, 0.6) {};

% Only video stalls annotation
\draw[thick, red!60, decorate,
  decoration={brace, amplitude=4pt}]
  (8.65, 2.65) -- (10.55, 2.65)
  node[midway, above=4pt, font=\scriptsize, text=red!60]
  {only video stalls};

\node[font=\scriptsize, text=green!50!black, anchor=west] at (12, 1.4)
  {unaffected};
\node[font=\scriptsize, text=green!50!black, anchor=west] at (12, 0.6)
  {unaffected};

% Legend
\node[font=\scriptsize, text=gray, anchor=north] at (6.25, -0.6) {%
  \tikz[baseline=-0.5ex]{\node[ok=blue!15, minimum width=0.35cm,
    minimum height=0.3cm] {};}~delivered \quad
  \tikz[baseline=-0.5ex]{\node[lost, minimum width=0.35cm,
    minimum height=0.3cm] {};}~lost \quad
  \tikz[baseline=-0.5ex]{\node[blocked, minimum width=0.35cm,
    minimum height=0.3cm] {};}~blocked};

\end{tikzpicture}
\caption{Head-of-line blocking: TCP vs.\ QUIC.  Left: when a TCP packet is
  lost, \emph{all} multiplexed streams stall until the lost packet is
  retransmitted.  Right: QUIC isolates the loss to the affected stream;
  audio and data continue flowing without interruption.}
\label{fig:hol-blocking-concept}
\end{figure}

\textbf{Built-in encryption.}  QUIC integrates TLS~1.3 (Transport Layer
Security, the same encryption standard used for HTTPS) directly into the
connection handshake.  A new QUIC connection is established in a single
network round trip (1-RTT), compared to TCP's three-way handshake followed
by a separate TLS handshake (typically 2--3 round trips).  For servers the
client has previously connected to, QUIC can establish a connection with
\emph{zero} round trips (0-RTT), sending application data in the very first
packet.

\textbf{Connection migration.}  Traditional TCP connections are identified
by a \emph{5-tuple}: source IP address, source port, destination IP address,
destination port, and protocol number.  When any of these change---for
example, when a phone switches from Wi-Fi to cellular---the TCP connection
breaks and must be re-established.  QUIC instead identifies connections by
\emph{connection IDs}, opaque tokens that survive network changes.  When a
device changes networks, QUIC validates the new path and continues the
session without interruption.  For a video call, this means seamless handoff
rather than a disconnect-and-reconnect.

\subsection{HTTP/3: HTTP over QUIC}

\textbf{HTTP/3 (RFC~9114)} is the latest version of the HTTP protocol that
powers the web.  It maps HTTP requests and responses onto QUIC instead of
TCP.  Each HTTP request/response pair occupies its own QUIC stream, so the
head-of-line blocking that affects HTTP/2 (which multiplexes over a single
TCP connection) is eliminated: a lost packet delays only the request it
belongs to, not every other request on the connection.

For the standards discussed in this document, HTTP/3 matters primarily
because it is the foundation on which WebTransport is built.  A
WebTransport session is established over an HTTP/3 connection, inheriting
all of QUIC's properties: independent streams, built-in encryption,
connection migration, and low-latency connection establishment.

\subsection{WebTransport: QUIC Capabilities for Browser Applications}

\textbf{WebTransport (W3C Working Draft + RFC~9297)} is a browser API that
gives JavaScript applications direct access to QUIC's transport capabilities
over an HTTP/3 connection.  To understand why this matters, consider the
alternative: \textbf{WebSocket}.

WebSocket, the current standard for bidirectional browser-server
communication, runs over TCP.  It provides a single ordered byte stream---
the same model as a regular TCP connection, with the same head-of-line
blocking problem.  If a WebSocket sends a burst of small messages and one
packet is lost, all subsequent messages stall until the lost one is
retransmitted.  Furthermore, every message sent over WebSocket is guaranteed
to arrive (TCP will retransmit until it does), even if the message has become
stale by the time it finally arrives.

WebTransport offers three capabilities that WebSocket cannot:

\begin{itemize}
  \item \textbf{Unreliable datagrams.}  An application can send a message
    that is delivered if possible but silently dropped if lost.  No
    retransmission, no stalling.  This is ideal for data that becomes stale
    quickly---game state updates, real-time cursor positions, live sensor
    readings---where receiving the \emph{latest} value matters more than
    receiving \emph{every} value.

  \item \textbf{Multiple independent reliable streams.}  An application can
    open many parallel streams, each with guaranteed in-order delivery, but
    without head-of-line blocking between them.  A stall on one stream does
    not affect the others.  This is useful for sending different types of
    data (e.g., chat messages, control signals, file transfers)
    simultaneously without interference.

  \item \textbf{Connection migration.}  Inherited from QUIC: the
    connection survives network changes such as Wi-Fi to cellular handoffs.
\end{itemize}

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  block/.style={draw, rounded corners=2pt, minimum height=0.55cm,
    align=center, font=\scriptsize, thick},
  heading/.style={font=\small\bfseries},
  arr/.style={->, thick, gray!50},
  lbl/.style={font=\tiny, text=gray},
]

% === WebSocket side (left) ===
\node[heading] at (2.0, 4.0) {WebSocket (TCP)};

\node[block, fill=orange!10, minimum width=3.6cm] (wsApp) at (2.0, 3.2)
  {Application};
\node[block, fill=gray!12, minimum width=3.6cm] (ws) at (2.0, 2.2)
  {WebSocket};
\node[block, fill=blue!8, minimum width=3.6cm] (tcp) at (2.0, 1.2)
  {TCP (single ordered stream)};
\node[block, fill=blue!8, minimum width=3.6cm] (tls) at (2.0, 0.2)
  {TLS 1.2/1.3};

\draw[arr] (wsApp) -- (ws);
\draw[arr] (ws) -- (tcp);
\draw[arr] (tcp) -- (tls);

% Limitations
\node[font=\tiny, text=red!50!black, anchor=north west, align=left]
  at (0.0, -0.5) {%
  $\times$ One stream (HOL blocking)\\
  $\times$ Reliable only (no datagrams)\\
  $\times$ No connection migration};

% === WebTransport side (right) ===
\node[heading] at (9.0, 4.0) {WebTransport (QUIC)};

\node[block, fill=orange!10, minimum width=5.0cm] (wtApp) at (9.0, 3.2)
  {Application};
\node[block, fill=green!10, minimum width=5.0cm] (wt2) at (9.0, 2.2)
  {WebTransport};

% Three capabilities
\node[block, fill=green!8, minimum width=1.4cm] (dg) at (7.2, 1.2)
  {Datagrams};
\node[block, fill=green!8, minimum width=1.4cm] (s1) at (9.0, 1.2)
  {Stream 1};
\node[block, fill=green!8, minimum width=1.4cm] (s2) at (10.8, 1.2)
  {Stream 2\ldots};

\node[block, fill=blue!8, minimum width=5.0cm] (quic2) at (9.0, 0.2)
  {QUIC + TLS 1.3 (independent streams)};

\draw[arr] (wtApp) -- (wt2);
\draw[arr] (wt2.south) -- ++(0, -0.25) -| (dg.north);
\draw[arr] (wt2.south) -- ++(0, -0.25) -| (s1.north);
\draw[arr] (wt2.south) -- ++(0, -0.25) -| (s2.north);
\draw[arr] (dg.south) -- ++(0, -0.15) -| ([xshift=-1.5cm]quic2.north);
\draw[arr] (s1) -- (quic2);
\draw[arr] (s2.south) -- ++(0, -0.15) -| ([xshift=1.5cm]quic2.north);

% Advantages
\node[font=\tiny, text=green!50!black, anchor=north west, align=left]
  at (6.3, -0.5) {%
  \checkmark{} Multiple independent streams\\
  \checkmark{} Unreliable datagrams (fire-and-forget)\\
  \checkmark{} Connection migration (Wi-Fi $\to$ cellular)};

% Label on datagrams
\node[font=\tiny, text=purple!60] at (7.2, 0.7) {unreliable};
\node[font=\tiny, text=blue!60] at (9.9, 0.7) {reliable};

% VS divider
\node[font=\large\bfseries, text=gray!30] at (5.5, 1.7) {vs.};

\end{tikzpicture}
\caption{WebSocket vs.\ WebTransport.  WebSocket provides a single ordered
  stream over TCP; a lost packet stalls everything.  WebTransport provides
  multiple independent streams \emph{and} unreliable datagrams over QUIC,
  with connection migration.}
\label{fig:ws-vs-wt}
\end{figure}

WebTransport is a \emph{client-server} protocol: the browser connects to a
server, not directly to another browser.  It complements WebRTC---where
WebRTC provides direct browser-to-browser media exchange, WebTransport
provides a modern browser-to-server data path suited to relayed media
architectures, game state synchronization, and real-time telemetry.

\subsection{WebCodecs: Direct Codec Access}

\textbf{WebCodecs (W3C Working Draft)} gives JavaScript code direct,
frame-by-frame access to the audio and video codecs built into the browser.

To understand why this matters, consider how browsers have traditionally
handled media encoding and decoding.  High-level APIs like
\texttt{MediaRecorder} (for recording) and \texttt{HTMLMediaElement} (for
playback) control the entire pipeline: the browser decides when to produce
keyframes, how to adjust bitrate, and what container format to use.  The
application gets a finished recording or a rendered video---but cannot
intervene at the level of individual frames.

WebCodecs breaks open this pipeline.  It provides four interfaces:

\begin{itemize}
  \item \texttt{VideoEncoder} / \texttt{VideoDecoder} --- encode or decode
    individual video frames, with per-frame control over keyframe insertion,
    target bitrate, and quality parameters.
  \item \texttt{AudioEncoder} / \texttt{AudioDecoder} --- the same for
    individual audio chunks.
\end{itemize}

These interfaces use the browser's built-in hardware-accelerated codec
implementations.  When dedicated encoding hardware is available (as on most
modern GPUs and mobile chips), WebCodecs uses it automatically; otherwise it
falls back to software encoding.  The application gets raw encoded bytes
without container overhead, and can process, transform, or transport them
however it chooses.

WebCodecs is deliberately \emph{transport-agnostic}---it handles encoding
and decoding, not delivery.  The encoded output can be sent via WebRTC (using
Encoded Transforms to insert custom processing into the
\texttt{RTCPeerConnection} pipeline), via WebTransport (packaging encoded
frames as datagrams or stream data), or via any other mechanism.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  block/.style={draw, rounded corners=2pt, minimum height=0.6cm,
    align=center, font=\scriptsize, thick},
  heading/.style={font=\small\bfseries},
  arr/.style={->, thick, gray!60},
  lbl/.style={font=\tiny, text=gray},
]

% === Traditional path (top) ===
\node[heading] at (0.8, 2.4) {Traditional};

\node[block, fill=blue!10, minimum width=1.6cm] (cam1) at (0.8, 1.6)
  {Camera};
\node[block, fill=gray!15, minimum width=3.2cm, minimum height=1.2cm]
  (mr) at (4.2, 1.6)
  {MediaRecorder\\{\tiny(browser controls everything)}};
\node[block, fill=orange!10, minimum width=2.0cm] (out1) at (7.8, 1.6)
  {Container file\\{\tiny(.webm, .mp4)}};

\draw[arr] (cam1) -- (mr);
\draw[arr] (mr) -- (out1);

\node[font=\tiny, text=red!50!black, anchor=north] at (4.2, 0.75)
  {No per-frame control};

% === WebCodecs path (bottom) ===
\node[heading] at (0.8, -0.2) {WebCodecs};

\node[block, fill=blue!10, minimum width=1.6cm] (cam2) at (0.8, -1.0)
  {Camera};
\node[block, fill=green!12, minimum width=2.0cm] (ve) at (3.5, -1.0)
  {VideoEncoder\\{\tiny(per-frame control)}};
\node[block, fill=purple!8, minimum width=1.6cm] (raw) at (6.2, -1.0)
  {Raw encoded\\frames};

% Multiple output paths
\node[block, fill=orange!8, minimum width=1.8cm] (rtc2) at (9.0, -0.2)
  {WebRTC};
\node[block, fill=orange!8, minimum width=1.8cm] (wt3) at (9.0, -1.0)
  {WebTransport};
\node[block, fill=orange!8, minimum width=1.8cm] (other) at (9.0, -1.8)
  {File / WASM / \ldots};

\draw[arr] (cam2) -- (ve);
\draw[arr] (ve) -- (raw);
\draw[arr] (raw.east) -- ++(0.3, 0) |- (rtc2.west);
\draw[arr] (raw) -- (wt3);
\draw[arr] (raw.east) -- ++(0.3, 0) |- (other.west);

\node[font=\tiny, text=green!50!black, anchor=north] at (6.5, -2.3)
  {Application chooses transport};

% Brace for HW accel
\draw[decorate, decoration={brace, amplitude=3pt, mirror},
  thick, blue!40] (2.4, -1.6) -- (4.6, -1.6)
  node[midway, below=4pt, font=\tiny, text=blue!50] {HW-accelerated};

\end{tikzpicture}
\caption{Traditional codec access vs.\ WebCodecs.  Top: \texttt{MediaRecorder}
  controls the entire pipeline; the application receives a finished container
  file.  Bottom: WebCodecs gives per-frame control over hardware-accelerated
  encoding, and the application decides how to transport the raw output.}
\label{fig:webcodecs-pipeline}
\end{figure}

\subsection{How the Standards Interact}

The standards described above can be summarized in four key relationships:

\begin{enumerate}
  \item \textbf{WebRTC is self-contained for peer-to-peer communication.}
    A single API call, \texttt{RTCPeerConnection}, bundles ICE (path
    discovery), DTLS-SRTP (encryption), RTP (media transport), and codec
    negotiation.  An application can build a complete video call using only
    WebRTC and an application-provided signaling channel, without touching
    any other standard discussed here.

  \item \textbf{QUIC $\to$ HTTP/3 $\to$ WebTransport is a layered
    client-server path.}  Each layer adds capability: QUIC provides
    multiplexed encrypted transport over UDP, HTTP/3 maps HTTP semantics
    onto it, and WebTransport exposes QUIC's streams and datagrams to
    JavaScript.  An application that needs browser-to-server real-time
    communication uses all three layers, but interacts only with the
    WebTransport API at the top.

  \item \textbf{WebCodecs bridges both paths.}  In a WebRTC context,
    WebCodecs (via Encoded Transforms) enables custom processing of media
    frames flowing through \texttt{RTCPeerConnection}---for example,
    applying end-to-end encryption inside the WebRTC pipeline.  In a
    WebTransport context, WebCodecs encodes and decodes media that the
    application transports itself, independent of WebRTC.

  \item \textbf{The ``unbundled'' alternative.}  Rather than using WebRTC's
    monolithic \texttt{RTCPeerConnection}, an application can compose
    WebTransport (for transport) + WebCodecs (for encoding/decoding) +
    WebAssembly (for custom processing) to build a fully custom media
    pipeline.  This trades WebRTC's built-in integration for explicit control
    over every decision: which codec, what packetization, how to handle
    congestion, what to retransmit.  This architecture is projected to become
    viable by 2029 (Sections~8--9).
\end{enumerate}

\section{WebRTC}

The WebRTC stack in the \texttt{standards/} directory comprises the W3C
WebRTC 1.0 specification and six IETF RFCs: JSEP (RFC~8829), ICE
(RFC~8445), the WebRTC security considerations (RFC~8826), the DTLS-based
security architecture (RFC~8827), RTP (RFC~3550), and SRTP (RFC~3711).

\subsection{Late 2021}

W3C WebRTC 1.0 achieved Recommendation status in January 2021, making it a
finalized standard. By late 2021, Chrome and Firefox offered full
implementations, and Edge inherited Chrome's support via its Chromium base.
Safari supported the core API but with significant limitations:

\begin{itemize}
  \item \textbf{Codec restrictions.} Safari supported only H.264 for video.
    Chrome and Firefox supported VP8, VP9, and H.264. This meant Safari could
    not participate in VP9-based calls and lacked access to VP9's more
    efficient bandwidth usage.
  \item \textbf{No simulcast or SVC.} Safari could not send multiple
    encodings of the same video stream at different resolutions (simulcast)
    or use Scalable Video Coding (SVC). Chrome and Firefox supported both,
    enabling adaptive quality based on network conditions.
  \item \textbf{RTCDataChannel issues on iOS.} The data channel did not work
    reliably on iOS Safari. Cross-browser data channel connections between
    Safari on iOS and Chrome or Firefox failed due to missing local ICE
    candidates on iOS.
  \item \textbf{SDP handling divergence.} Chrome was in the process of
    migrating from Plan~B to Unified Plan SDP semantics. Chrome~M89
    (February 2021) added deprecation warnings; Chrome~M93 (August 2021)
    removed Plan~B with a reverse origin trial option. Firefox and Safari
    already used Unified Plan exclusively. During this transition, applications
    had to accommodate both SDP formats.
  \item \textbf{Insertable Streams.} Chrome~M94 introduced the
    Insertable Streams API, enabling JavaScript to access and transform
    encoded media frames in the WebRTC pipeline. This was essential for
    end-to-end encryption of media. Firefox and Safari did not implement it.
  \item \textbf{Stats reporting inconsistencies.} Firefox was missing
    \texttt{voice\-Activity\-Flag} on video track stats. Safari was missing
    \texttt{rtpTimestamp} and \texttt{voiceActivityFlag}. Applications could
    not reliably detect voice activity or obtain precise timing information
    across browsers.
\end{itemize}

On the IETF side, the supporting protocols were in the following state:
\begin{itemize}
  \item \textbf{JSEP (RFC~8829)} was published in January 2021 alongside the
    W3C Recommendation. All browsers implemented the JSEP signaling
    state machine through \texttt{RTCPeerConnection}, but diverged in areas
    like BUNDLE handling of multiple media streams.
  \item \textbf{ICE (RFC~8445)} was implemented in all browsers. Trickle ICE
    (RFC~8838), enabling incremental candidate exchange rather than waiting for
    full gathering, was published in January 2021 but not yet mandatory.
  \item \textbf{DTLS-SRTP (RFCs~8827, 3711)} was mandatory across all
    browsers. All media was encrypted. However, browsers differed in DTLS
    certificate algorithms (Chrome~52+ used ECDSA; others varied), and
    privacy identifier reset mechanisms were inconsistent.
  \item \textbf{Security (RFC~8826)} --- all browsers enforced
    camera/microphone permission prompts and encrypted all media. However,
    protections against IP address leakage were inconsistent; not all browsers
    masked local IP addresses via mDNS candidates when media permissions had
    not been granted.
\end{itemize}

\subsection{March 2026}

The W3C updated the WebRTC 1.0 Recommendation in October 2024 and again in
2025, incorporating amendments based on implementation experience.

\begin{itemize}
  \item \textbf{Unified Plan is universal.} Plan~B SDP is completely removed
    from all browsers. Applications no longer need to accommodate both formats.
  \item \textbf{Safari gained VP9.} Safari~15 and later support the VP9 codec
    for WebRTC, enabling Safari to participate in VP9-based calls and benefit
    from VP9's bandwidth efficiency. Safari still prefers H.264 for hardware
    acceleration, and its SVC support remains limited compared to
    Chrome and Firefox.
  \item \textbf{iOS WebRTC matured.} Starting with iOS~14.3, WebRTC works
    reliably across all browsers on iOS. RTCDataChannel connections between
    iOS Safari and desktop Chrome or Firefox are reliable.
  \item \textbf{JSEP updated to RFC~9429.} Published in April 2024, RFC~9429
    obsoleted RFC~8829, addressing implementation divergence --- particularly
    in BUNDLE handling and bundle-only \texttt{m=} sections.
  \item \textbf{Trickle ICE is mandatory.} All browsers implement RFC~8838,
    enabling faster connection establishment through incremental candidate
    exchange.
  \item \textbf{mDNS privacy protections standardized.} Chrome, Firefox, and
    Safari all mask local IP addresses using mDNS candidates when
    camera/microphone permissions have not been granted, implementing the
    privacy considerations of RFC~8826.
  \item \textbf{Encoded Transforms remain Chrome-only.} The feature
    evolved from ``Insertable Streams'' to \texttt{RTCRtp\-Script\-Transform}
    (WebRTC Encoded Transform), but Firefox and Safari still do not
    implement it. End-to-end encryption of encoded media frames in the
    WebRTC pipeline remains a Chrome/Chromium-only capability.
\end{itemize}

\section{QUIC}

RFC~9000 defines the QUIC transport protocol: a UDP-based, multiplexed,
encrypted transport with built-in TLS~1.3, independent stream flow control,
and connection migration across network changes.

\subsection{Late 2021}

RFC~9000 was published in May 2021. Browser support was uneven:

\begin{itemize}
  \item \textbf{Chrome} had QUIC enabled by default, having been an early
    adopter (Google developed the original QUIC protocol).
  \item \textbf{Firefox~88} (April 2021) enabled QUIC by default.
  \item \textbf{Safari} had experimental QUIC support (added in Safari
    Technology Preview~104, April 2020), but it required manual opt-in
    through experimental settings. It was not available to ordinary users.
\end{itemize}

In practice, a browser connecting to a QUIC-capable server would benefit from
reduced connection latency (0-RTT or 1-RTT handshakes vs.\ TCP's multi-round
handshake plus TLS), elimination of head-of-line blocking across multiplexed
streams, and connection survival across network changes. But with Safari
requiring manual opt-in, these benefits were unavailable to most Safari users.

\subsection{March 2026}

QUIC is enabled by default in all major browsers:
\begin{itemize}
  \item Chrome, Edge, and Firefox continue default support.
  \item Safari~16 (2022) enabled QUIC by default, bringing universal coverage.
\end{itemize}

Every major browser can now establish QUIC connections transparently. The
capabilities defined by RFC~9000 --- low-latency connection establishment,
multiplexed streams without head-of-line blocking, built-in encryption, and
connection migration --- are universally available to web applications without
any user or developer opt-in.

\section{HTTP/3}

RFC~9114 maps HTTP semantics over QUIC, inheriting QUIC's transport
properties: multiplexed streams without head-of-line blocking, lower
connection latency, and connection migration.

\subsection{Late 2021}

RFC~9114 had not yet been published (it was finalized in June 2022). Browsers
were shipping pre-standard HTTP/3 implementations based on draft
specifications:

\begin{itemize}
  \item \textbf{Chrome} had experimental HTTP/3 support enabled by default
    since April 2020.
  \item \textbf{Firefox~88} (April 2021) enabled HTTP/3 by default.
  \item \textbf{Safari} had HTTP/3 support behind an experimental feature
    flag, disabled by default.
\end{itemize}

Where supported, browsers could negotiate HTTP/3 with compatible servers,
gaining per-stream independence (a stalled stream would not block other
streams, unlike HTTP/2 over TCP). However, the lack of a finalized RFC meant
implementations were based on draft specifications that could diverge, and
Safari's experimental-only status limited universal availability.

\subsection{March 2026}

With RFC~9114 finalized in June 2022, all major browsers now support HTTP/3
by default:
\begin{itemize}
  \item Chrome, Edge, and Firefox have supported HTTP/3 by default for
    several years.
  \item Safari enabled HTTP/3 by default starting with Safari~16 (2022).
  \item Over 92\% of browser market share supports HTTP/3.
\end{itemize}

The capabilities defined by RFC~9114 --- HTTP request/response multiplexing
without head-of-line blocking, lower latency connection establishment via
QUIC, and connection resilience across network changes --- are now universally
available. Any browser connecting to an HTTP/3-capable server automatically
benefits from these properties.

\section{WebTransport}

The WebTransport specification defines browser APIs for bidirectional
client-server communication over HTTP/3, supporting both reliable streams and
unreliable datagrams. RFC~9297 defines the HTTP datagram and capsule protocol
mechanisms that underpin it.

\subsection{Late 2021}

WebTransport was in a very early state:

\begin{itemize}
  \item \textbf{Chrome} had experimental support but it was disabled in
    Chrome~M91 through M94. The API was not usable in production.
  \item \textbf{Firefox} had no support whatsoever.
  \item \textbf{Safari} had no support whatsoever.
  \item RFC~9297 had not yet been published (August 2022).
  \item The W3C specification was an early Working Draft.
\end{itemize}

In late 2021, WebTransport was effectively unavailable in any browser for
production use. The capabilities it defines --- unreliable datagrams from
browser to server, multiple independent reliable streams, and connection
migration --- could not be used by web applications.

\subsection{March 2026}

WebTransport has achieved broad but not universal support:

\begin{itemize}
  \item \textbf{Chrome~97+} and \textbf{Edge~98+} support WebTransport
    (stable since late 2021 / early 2022).
  \item \textbf{Firefox~114+} (2023) supports WebTransport.
  \item \textbf{Safari does not support WebTransport.}
  \item Approximately 82\% of browser market share has WebTransport support.
\end{itemize}

For browsers that support it, WebTransport enables capabilities that were
previously impossible without WebRTC's peer-to-peer architecture:
\begin{itemize}
  \item \textbf{Unreliable datagrams} from browser to server, suitable for
    latency-sensitive data where retransmission is undesirable (e.g., game
    state updates, real-time sensor data).
  \item \textbf{Multiple independent reliable streams} without head-of-line
    blocking between them, enabling parallel data flows over a single
    connection.
  \item \textbf{Connection migration} across network changes, inherited from
    QUIC.
\end{itemize}

Safari's absence means applications using WebTransport must provide fallback
mechanisms (typically WebSocket) for Safari users, or accept that these
capabilities are unavailable on Apple platforms.

\section{WebCodecs}

The WebCodecs specification provides JavaScript interfaces for direct access
to the browser's built-in audio, video, and image codecs, enabling
frame-level encoding and decoding outside of higher-level APIs like
\texttt{MediaRecorder} or \texttt{HTMLMediaElement}.

\subsection{Late 2021}

WebCodecs had just reached initial availability:

\begin{itemize}
  \item \textbf{Chrome~94} (September 2021) shipped WebCodecs as a stable
    feature, graduating from origin trial. Chrome~92 and 93 had offered it
    as an origin trial with breaking changes between versions.
  \item \textbf{Firefox} had no support.
  \item \textbf{Safari} had no support.
  \item The W3C specification was a First Public Working Draft (2021).
\end{itemize}

In Chrome~94, developers could for the first time:
\begin{itemize}
  \item Encode and decode individual video frames using
    \texttt{VideoEncoder} and \texttt{VideoDecoder} with the browser's
    hardware-accelerated codecs, without being constrained by
    \texttt{MediaRecorder}'s real-time recording model.
  \item Encode and decode individual audio chunks
    (\texttt{AudioEncoder}/\texttt{AudioDecoder}).
  \item Perform faster-than-realtime encoding (previously impossible via
    \texttt{MediaRecorder}).
  \item Control per-frame keyframe insertion and quality parameters.
  \item Obtain raw encoded bytes without container overhead.
\end{itemize}

However, these capabilities were Chrome-only. Firefox and Safari users had no
access to low-level codec APIs; they were limited to the high-level
\texttt{MediaRecorder} and \texttt{HTMLMediaElement} interfaces.

\subsection{March 2026}

WebCodecs has achieved broad cross-browser support:

\begin{itemize}
  \item \textbf{Chrome/Edge} --- supported since version 94 (2021).
  \item \textbf{Firefox~130+} (September 2024) --- shipped approximately three
    years after Chrome.
  \item \textbf{Safari~26+} (2024) --- full support. Earlier Safari versions
    (16.4+) had partial video-only support.
  \item Approximately 94\% of browser market share supports WebCodecs.
  \item The specification remains a W3C Working Draft (updated January 2026),
    not yet a Recommendation.
\end{itemize}

With cross-browser support, all major browsers can now perform frame-level
codec operations using hardware acceleration. This enables web applications
to build custom media pipelines --- video editors, live transcoders, custom
streaming clients --- that were previously possible only through WebAssembly
codec implementations or native applications.

\section{Looking Ahead: 2028--2029 (Two to Three Years)}

The following projections are grounded in active W3C charter milestones, IETF
draft publication timelines, and formal browser vendor implementation signals.

\subsection{WebCodecs Reaches W3C Recommendation}

The W3C Media Working Group charter targets WebCodecs Candidate
Recommendation in Q1~2026 and full Recommendation by Q4~2026. By 2028, the
specification should be a mature, finalized standard. This means the
frame-level codec access that was Chrome-only in 2021 and cross-browser by
2026 will have a stable, locked-down API surface that browser vendors commit
to maintaining long-term. The WebCodecs Codec Registry will likely include
formal AV1 registration entries alongside existing H.264, VP8, and VP9
entries.

\subsection{WebTransport Achieves Universal Browser Support}

Apple committed WebTransport as a formal Interop~2026 priority, accounting
for 20\% of the interoperability score. Safari already has experimental
WebTransport support behind a flag in iOS~18. Given this commitment, Safari
production support is expected by late 2026 or early 2027, with the W3C
specification reaching Recommendation status around the same time. By 2028,
WebTransport should be universally available across all major browsers ---
closing the last significant gap in the stack. Applications will be able to
rely on browser-to-server unreliable datagrams and independent reliable
streams without Safari fallbacks.

\subsection{WebRTC Encoded Transforms Become Cross-Browser}

The \texttt{RTCRtpScriptTransform} API (WebRTC Encoded Transform) has a
Working Draft dated February 2026 and is approaching Candidate
Recommendation. Safari and Firefox have implemented it natively; Chrome
continues support via its earlier Insertable Streams lineage. By 2028, this
API should be standardized and available across all major browsers, enabling
cross-browser end-to-end encryption of media frames within the WebRTC
pipeline --- a capability that was Chrome-only from 2021 through 2026.

\subsection{AV1 Becomes a Practical WebRTC Codec}

Firefox enabled AV1 for WebRTC by default in 2025. Chrome has supported it
for several years. By 2028, AV1 hardware encoding support in consumer devices
will have expanded, reducing the CPU cost that currently limits its adoption.
AV1's superior compression efficiency (roughly 30\% better than VP9 at
equivalent quality) will make it a practical default codec for WebRTC calls
in Chrome and Firefox. Safari's adoption remains uncertain, as Apple
historically favors H.264 with hardware acceleration.

\subsection{Media over QUIC Reaches RFC Status}

The IETF Media over QUIC (MoQ) transport draft
(\texttt{draft-ietf-moq-transport}) is advancing rapidly, with version~15
published in October 2025. The protocol enables subscription-based media
publishing and consuming over QUIC, designed for low-latency distribution
with CDN support. By 2028, MoQ should have reached RFC status, along with
companion specifications for media container formats (Low Overhead Media
Container) and streaming formats. Browser integration will likely occur
through JavaScript libraries built on WebTransport and WebCodecs rather than
through dedicated browser APIs --- the component APIs are already in place to
support this.

\subsection{RTP over QUIC Reaches RFC Status}

The IETF RTP over QUIC (RoQ) draft (\texttt{draft-ietf-avtcore-rtp-over-quic})
maps RTP and RTCP packets onto QUIC streams and datagrams. By 2028, this
should reach RFC status, enabling WebRTC backends to transport media over
QUIC instead of the traditional UDP+DTLS-SRTP path. Browser-facing
integration would occur through SDP negotiation rather than new JavaScript
APIs, potentially enabling WebRTC calls to benefit from QUIC's congestion
control and connection migration at the transport layer.

\subsection{Hardware and Network Context}

The standards projections above do not exist in a vacuum. By 2028, two
underlying trends amplify what these standards make practically possible.

\textbf{CPU and hardware acceleration.} AV1 encoding is computationally
expensive --- in 2026, real-time AV1 encoding for a WebRTC call can consume
an entire CPU core on mid-range hardware, which is why VP9 and H.264 remain
the practical defaults despite AV1's superior compression. By 2028, dedicated
AV1 hardware encoders will be standard in new laptops, phones, and tablets
(Apple's M-series chips, Qualcomm Snapdragon, Intel Arc, and AMD RDNA
already include or are adding AV1 encode blocks). This shifts AV1 from
``technically supported but too expensive'' to ``hardware-accelerated and
free.'' The WebCodecs API, which exposes hardware-accelerated codecs to
JavaScript, directly benefits: an application calling
\texttt{VideoEncoder} with AV1 in 2028 will hit dedicated silicon rather than
a software encoder, making frame-level AV1 processing via WebCodecs
practical for real-time use cases.

Similarly, the unbundled architecture --- composing WebTransport, WebCodecs,
and WebAssembly --- imposes CPU overhead from JavaScript orchestration and
WebAssembly processing that would have been prohibitive on 2021 hardware.
By 2028, the roughly 40--50\% improvement in single-thread CPU performance
since 2021 (compounding at approximately 8--10\% annually) makes this
overhead manageable for real-time media pipelines.

\textbf{Network bandwidth.} Global median fixed broadband speeds have
approximately doubled from 2021 to 2026 (from roughly 60~Mbps to over
120~Mbps in developed markets), and 5G mobile networks have expanded
substantially. By 2028, continued growth in fixed and mobile bandwidth
means that the efficiency gains from AV1 and QUIC compound with raw
bandwidth increases. A WebRTC call using AV1 over QUIC on a 2028 connection
will have both a more efficient codec (30\% less bandwidth than VP9) and
more available bandwidth, enabling higher resolution and frame rates within
the same network cost envelope. QUIC's connection migration (RFC~9000)
becomes more impactful as mobile usage grows --- a video call that seamlessly
survives a Wi-Fi to 5G handoff requires both the protocol support (already
standardized) and sufficient mobile bandwidth to sustain the call (increasingly
available).

The net effect is that the standards reaching maturity by 2028 will land on
hardware and networks that can fully exploit them, rather than being
bottlenecked by the infrastructure they run on.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  item/.style={draw, rounded corners=3pt, minimum height=0.85cm,
    align=center, font=\footnotesize, fill=white},
  arr/.style={->, thick, gray!60},
  metriclabel/.style={font=\scriptsize\itshape, text=black!60},
]

% --- Layer background bands (no border, distinct colors) ---
\fill[blue!8, rounded corners=4pt]
  (-5.5, -0.65) rectangle (6.7, 0.65);
\fill[green!10, rounded corners=4pt]
  (-5.5, 1.55) rectangle (6.7, 2.85);
\fill[orange!10, rounded corners=4pt]
  (-5.5, 3.75) rectangle (6.7, 5.05);

% --- Infrastructure items ---
\node[item, fill=blue!15] (hw) at (-3.2, 0)
  {AV1 HW encode\\M-series / Snapdragon / Arc};
\node[item, fill=blue!15] (cpu) at (0.7, 0)
  {CPU +40--50\%\\vs.\ 2021};
\node[item, fill=blue!15] (net) at (4.4, 0)
  {Broadband ${\sim}$150~Mbps\\5G expanding};

% --- Standards items ---
\node[item, fill=green!18] (wc) at (-4.2, 2.2)
  {WebCodecs\\(W3C Rec)};
\node[item, fill=green!18] (wt) at (-1.5, 2.2)
  {WebTransport\\(universal)};
\node[item, fill=green!18] (av1) at (1.2, 2.2)
  {AV1 in\\WebRTC};
\node[item, fill=green!18] (quic) at (3.5, 2.2)
  {QUIC / HTTP/3\\(RFC 9000/9114)};
\node[item, fill=green!18] (et) at (5.8, 2.2)
  {Encoded\\Transforms};

% --- User experience items ---
\node[item, fill=orange!18] (hd) at (-3.8, 4.4)
  {1080p / 4K\\video calls};
\node[item, fill=orange!18] (handoff) at (-1.0, 4.4)
  {Seamless\\Wi-Fi $\to$ 5G};
\node[item, fill=orange!18] (e2ee) at (1.9, 4.4)
  {Cross-browser\\E2E encryption};
\node[item, fill=orange!18] (custom) at (4.8, 4.4)
  {Custom media\\pipelines};

% --- Arrows: infrastructure -> standards ---
\draw[arr] (hw.north) -- (wc.south);
\draw[arr] (hw.north) -- (av1.south);
\draw[arr] (cpu.north) -- (wt.south);
\draw[arr] (cpu.north) -- (wc.south);
\draw[arr] (net.north) -- (quic.south);
\draw[arr] (net.north) -- (av1.south);

% --- Arrows: standards -> user experience ---
\draw[arr] (wc.north) -- (hd.south);
\draw[arr] (wc.north) -- (custom.south);
\draw[arr] (av1.north) -- (hd.south);
\draw[arr] (wt.north) -- (custom.south);
\draw[arr] (wt.north) -- (handoff.south);
\draw[arr] (quic.north) -- (handoff.south);
\draw[arr] (et.north) -- (e2ee.south);

% --- Legend (below diagram, horizontal) ---
\node[anchor=north, font=\footnotesize] at (0.5, -0.9) {%
  \tikz[baseline=-0.5ex]{\fill[orange!18] (0,0) rectangle (0.35,0.35);}
  ~User Experience \qquad
  \tikz[baseline=-0.5ex]{\fill[green!18] (0,0) rectangle (0.35,0.35);}
  ~Standards \qquad
  \tikz[baseline=-0.5ex]{\fill[blue!15] (0,0) rectangle (0.35,0.35);}
  ~Infrastructure};

% --- Measurable effects ---
\node[metriclabel, anchor=north, align=center] at (0.5, -1.5)
  {\textbf{Measurable effects:} ~30\% less bandwidth
   ~$\cdot$~ seamless handoff ~$\cdot$~
   E2EE in all browsers ~$\cdot$~ sub-second latency};

\end{tikzpicture}
\caption{By 2028: infrastructure enables standards to deliver measurable
  user experience gains. Arrows show enabling relationships.}
\label{fig:2028-stack}
\end{figure}

\section{Looking Ahead: 2029--2031 (Three to Five Years)}

These projections extend further from current standardization activity and
carry correspondingly more uncertainty, but follow from the trajectory
established by the standards work described above.

\subsection{The ``Unbundled'' Architecture Becomes Viable}

By 2029, all the component APIs --- WebTransport (universal), WebCodecs
(Recommendation), and RTCRtpScriptTransform (standardized) --- will have been
stable and cross-browser for two or more years. This makes the ``WebRTC
Unbundled'' architecture practically viable: applications can compose
WebTransport for data transport, WebCodecs for encoding and decoding, and
WebAssembly for custom processing, bypassing the monolithic
\texttt{RTCPeerConnection} API entirely.

This matters because the unbundled approach gives applications control over
decisions that WebRTC currently makes internally: which codec to use, how to
packetize, what transport to use, and how to handle congestion. Use cases
that are poorly served by WebRTC's built-in assumptions --- cloud gaming
requiring ultra-low latency with acceptable loss, scientific visualization
requiring custom frame processing, or broadcast distribution requiring
one-to-many fan-out --- can be built from these composable primitives.

Traditional WebRTC via \texttt{RTCPeerConnection} will persist for general
video conferencing, where its built-in codec negotiation, ICE connectivity,
and DTLS-SRTP security provide a well-integrated default. The unbundled
approach is not a replacement but an alternative for applications with
specialized requirements.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
  block/.style={draw, rounded corners=3pt, minimum height=0.8cm,
    align=center, font=\footnotesize, thick},
  monolith/.style={block, fill=gray!15, minimum width=5cm},
  unbundled/.style={block, fill=green!10, minimum width=2.6cm},
  label/.style={font=\small\bfseries, anchor=south},
  arr/.style={->, thick, gray!60},
  brace/.style={decorate, decoration={brace, amplitude=6pt, mirror}},
]

% --- Monolithic side (left) ---
\node[label] at (-3.5, 5.5) {Monolithic (WebRTC 1.0)};

\node[monolith] (app1) at (-3.5, 4.5) {Application};
\node[monolith, fill=gray!25] (rtc) at (-3.5, 3.3) {\texttt{RTCPeerConnection}};
\node[monolith, fill=gray!20, minimum height=1.6cm] (internals) at (-3.5, 1.7)
  {Codec negotiation\\ICE / DTLS-SRTP\\RTP / RTCP\\Congestion control};
\node[monolith] (udp1) at (-3.5, 0.3) {UDP};

\draw[arr] (app1) -- (rtc);
\draw[arr] (rtc) -- (internals);
\draw[arr] (internals) -- (udp1);

% fixed / chosen by browser label
\node[font=\scriptsize\itshape, text=gray, anchor=west] at (-0.7, 1.7) {\parbox{1.1cm}{fixed by\\browser}};

% --- Unbundled side (right) ---
\node[label] at (5.0, 5.5) {Unbundled (2029+)};

\node[unbundled, minimum width=7.2cm] (app2) at (5.0, 4.8) {Application};

\node[unbundled, fill=orange!12, minimum width=2.2cm] (wcodec) at (2.5, 3.1) {WebCodecs};
\node[unbundled, fill=cyan!12, minimum width=2.2cm] (wasm) at (5.0, 3.1) {WebAssembly};
\node[unbundled, fill=orange!12, minimum width=2.2cm] (et2) at (7.5, 3.1) {E2EE};

\node[unbundled, fill=blue!10, minimum width=2.8cm] (wtrans) at (3.4, 1.8) {WebTransport};
\node[unbundled, fill=blue!10, minimum width=2.8cm] (moq) at (6.6, 1.8) {MoQ (JS lib)};

\node[unbundled, fill=blue!8, minimum width=7.2cm] (quicl) at (5.0, 0.6) {QUIC / HTTP/3};

\draw[arr] (app2.south) -- ++(0,-0.55) -| (wcodec.north);
\draw[arr] (app2.south) -- ++(0,-0.55) -| (wasm.north);
\draw[arr] (app2.south) -- ++(0,-0.55) -| (et2.north);
\draw[arr] (wcodec) -- (wtrans);
\draw[arr] (wasm) -- (moq);
\draw[arr] (et2.south) -- (moq.north);
\draw[arr] (wtrans) -- (quicl);
\draw[arr] (moq) -- (quicl);

% chosen by application label
\node[font=\scriptsize\itshape, text=green!50!black, anchor=west,
  rotate=90] at (8.9, 2.0) {chosen by app};

% --- "vs" divider ---
\node[font=\large\bfseries, text=gray!40] at (0.5, 2.5) {vs.};

\end{tikzpicture}
\caption{Monolithic WebRTC (left) vs.\ the unbundled architecture (right).
  The unbundled approach gives applications control over codec, transport,
  encryption, and congestion decisions that \texttt{RTCPeerConnection}
  handles internally.}
\label{fig:unbundled}
\end{figure}

\subsection{MoQ Enables Standards-Based Low-Latency Streaming}

With MoQ at RFC status and WebTransport universally available, a
standards-based path exists for low-latency media streaming from server to
browser that does not require WebRTC. A MoQ JavaScript library running on
WebTransport and using WebCodecs for decoding could deliver sub-second
latency streaming with CDN fan-out --- combining the latency characteristics
previously unique to WebRTC with the scalability of traditional CDN
architectures. This creates a standards-based alternative to proprietary
low-latency streaming protocols.

\subsection{AV1 as the Default WebRTC Codec}

By 2030, hardware AV1 encoding will be standard in consumer devices
(laptops, phones, tablets). The CPU cost barrier that limits AV1 adoption in
2026 will have largely disappeared. AV1's compression advantage over VP9 and
H.264 makes it the natural default for WebRTC, reducing bandwidth
requirements for equivalent video quality. Whether Safari adopts AV1 for
WebRTC depends on Apple's hardware acceleration decisions, but the broader
ecosystem pressure from Chrome and Firefox --- which together represent the
majority of browser usage --- will make AV1 the dominant codec for new
WebRTC applications.

\subsection{QUIC v2 Transparent Adoption}

QUIC~v2 (RFC~9369, published May 2023) is identical to QUIC~v1 in capability
but uses different wire constants to resist middlebox ossification. By 2030,
browsers will have transparently migrated to QUIC~v2 for new connections.
This is invisible to applications --- no new capabilities are exposed --- but
ensures the QUIC transport layer remains resilient against network
intermediaries that might otherwise interfere with the protocol based on
recognizing v1 byte patterns.

\subsection{No WebRTC 2.0}

The W3C WebRTC Working Group charter (extending to September 2026) adds new
capabilities through extension specifications rather than monolithic version
bumps. This pattern is expected to continue. There will be no ``WebRTC~2.0''
in the traditional sense. Instead, the API surface will grow incrementally
through standardized extensions --- Encoded Transforms, SFrame for
end-to-end encryption in group calls, and future extensions not yet drafted
--- while the core \texttt{RTCPeerConnection} API remains stable.

\subsection{Hardware and Network Context}

By 2030, the hardware and network landscape further amplifies the practical
impact of these matured standards.

\textbf{CPU and hardware acceleration.} By 2030, single-thread CPU
performance will have improved roughly 60--80\% relative to 2021 levels,
and AV1 hardware encode/decode will be ubiquitous --- present not just in
flagship devices but in budget laptops and mid-range phones. This has two
consequences for the standards discussed here. First, AV1 becomes the
natural default codec for WebRTC: the hardware acceleration eliminates the
CPU cost barrier, and WebCodecs applications calling \texttt{VideoEncoder}
with AV1 will consistently hit dedicated silicon across the device spectrum,
not just on high-end hardware. Second, the unbundled architecture ---
WebTransport plus WebCodecs plus WebAssembly --- becomes practical even on
resource-constrained devices. The JavaScript and WebAssembly orchestration
overhead that would strain a 2021 mobile processor is routine on a 2030
device, making custom real-time media pipelines viable on phones and tablets,
not just desktop machines.

\textbf{Network bandwidth.} If current trends hold, median fixed broadband
speeds in developed markets will reach 200--300~Mbps by 2030, with 5G mobile
networks providing reliable 100+~Mbps in urban areas. At these speeds,
bandwidth is no longer the binding constraint for most real-time
communication scenarios. The practical consequence is a shift in what
matters: codec efficiency (AV1 over H.264) becomes less about fitting within
tight bandwidth limits and more about enabling higher quality --- 4K video
conferencing, high-frame-rate screen sharing, or multi-stream layouts ---
without saturating the connection. QUIC's multiplexed streams (RFC~9000)
and HTTP/3's head-of-line blocking elimination (RFC~9114) matter most when
multiple parallel flows compete for bandwidth, which is precisely the
scenario enabled by abundant bandwidth and ambitious applications.

MoQ-based streaming benefits particularly from this environment. A MoQ
library on WebTransport using WebCodecs to decode AV1 can deliver
high-bitrate, low-latency streams that would have been impractical in 2021
not because the standards didn't exist, but because the combination of codec
cost, network capacity, and API availability wasn't there. By 2030, all
three constraints are resolved simultaneously: the standards are finalized
and universal, the hardware accelerates the codecs, and the networks carry
the bits.

\textbf{The compounding effect.} The key observation is that standards
maturation, hardware improvement, and network growth are complementary, not
independent. A standard that reaches browser universality on hardware that
can't run it efficiently, or over networks that can't carry its output,
delivers theoretical capability rather than practical impact. By the
2029--2031 timeframe, these three curves converge: every standard in this
document will be finalized and universally implemented, running on hardware
with dedicated acceleration for the codecs they specify, over networks with
bandwidth to spare. This convergence is what makes the 3--5 year outlook
qualitatively different from a simple extrapolation of today's capabilities.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  >=Stealth,
]

% --- Axes ---
\draw[->, thick] (0, 0) -- (12.5, 0) node[anchor=north, font=\small] {};
\draw[->, thick] (0, 0) -- (0, 5.5) node[anchor=east, font=\small, rotate=90,
  yshift=6pt] {Practical capability};

% Year labels
\foreach \x/\yr in {0.5/2021, 3.5/2024, 6.5/2026, 9/2028, 11.5/2031} {
  \draw (\x, -0.1) -- (\x, 0.1);
  \node[font=\footnotesize, anchor=north] at (\x, -0.15) {\yr};
}

% --- Standards maturity curve (green) ---
\draw[very thick, green!60!black]
  (0.5, 0.8) .. controls (2.5, 1.2) and (4, 2.5) ..
  (6.5, 3.8) .. controls (8, 4.4) and (9.5, 4.7) ..
  (11.5, 4.9);
\node[font=\scriptsize, green!60!black, anchor=south west] at (10.2, 4.85)
  {Standards};

% --- Hardware capability curve (blue) ---
\draw[very thick, blue!70]
  (0.5, 1.5) .. controls (2.5, 1.8) and (4.5, 2.3) ..
  (6.5, 2.9) .. controls (8, 3.5) and (9.5, 4.0) ..
  (11.5, 4.6);
\node[font=\scriptsize, blue!70, anchor=south west] at (10.2, 4.4)
  {Hardware};

% --- Network bandwidth curve (orange) ---
\draw[very thick, orange!80!black]
  (0.5, 1.2) .. controls (2.5, 1.6) and (4.5, 2.1) ..
  (6.5, 2.7) .. controls (8, 3.2) and (9.5, 3.7) ..
  (11.5, 4.3);
\node[font=\scriptsize, orange!80!black, anchor=south west] at (10.2, 3.95)
  {Network};

% --- Convergence zone ---
\fill[gray!12, rounded corners=5pt] (8.5, 3.4) rectangle (11.8, 5.2);
\node[font=\scriptsize\itshape, text=black!50, align=center] at (10.15, 3.6)
  {convergence zone};

% --- Key milestones (annotation dots) ---
% WebRTC 1.0 Rec
\fill[green!60!black] (0.5, 0.8) circle (2.5pt);
\node[font=\tiny, anchor=south, green!60!black, yshift=2pt] at (0.5, 0.8)
  {WebRTC Rec};

% QUIC RFC
\fill[green!60!black] (1.3, 1.0) circle (2.5pt);
\node[font=\tiny, anchor=north west, green!60!black] at (1.0, 0.85)
  {QUIC RFC};

% HTTP/3 RFC
\fill[green!60!black] (3.0, 1.8) circle (2.5pt);
\node[font=\tiny, anchor=south, green!60!black, yshift=2pt] at (3.0, 1.8)
  {HTTP/3 RFC};

% WebCodecs cross-browser
\fill[green!60!black] (5.5, 3.2) circle (2.5pt);
\node[font=\tiny, anchor=south east, green!60!black] at (5.6, 3.2)
  {WebCodecs x-browser};

% WebTransport universal (projected)
\fill[green!60!black] (8.0, 4.2) circle (2.5pt);
\node[font=\tiny, anchor=south, green!60!black, yshift=2pt] at (8.0, 4.2)
  {WebTransport universal};

% AV1 HW ubiquitous
\fill[blue!70] (9.5, 4.05) circle (2.5pt);
\node[font=\tiny, anchor=north, blue!70, yshift=-2pt] at (9.5, 4.05)
  {AV1 HW ubiquitous};

\end{tikzpicture}
\caption{Convergence of standards maturity, hardware capability, and network
  bandwidth from 2021 to 2031. The shaded region marks where all three curves
  reach levels sufficient to fully exploit the standards discussed in this
  document.}
\label{fig:convergence}
\end{figure}

\section{Summary}

\begin{table}[ht]
\centering
\small
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{lXXX}
\toprule
\textbf{Standard} & \textbf{Late 2021} & \textbf{March 2026} & \textbf{2028--2031 Projection} \\
\midrule
WebRTC 1.0 &
  W3C Rec (Jan 2021). Chrome/Firefox full support. Safari limited: H.264 only, no simulcast, iOS DataChannel broken. Plan~B still in Chrome. &
  Updated Rec (2024--25). Unified Plan universal. Safari has VP9. iOS reliable. JSEP updated to RFC~9429. Encoded Transforms Chrome-only. &
  Encoded Transforms cross-browser. AV1 default codec (HW-accelerated). SFrame for group E2EE. No ``WebRTC 2.0'' --- incremental extensions. \\
\midrule
QUIC &
  RFC~9000 (May 2021). Chrome/Firefox default. Safari experimental, manual opt-in. &
  Default in all browsers (Safari~16+, 2022). &
  QUIC~v2 (RFC~9369) transparent migration. RTP over QUIC (RoQ) at RFC status. \\
\midrule
HTTP/3 &
  Pre-RFC draft implementations. Chrome/Firefox default. Safari behind flag. &
  RFC~9114 (June 2022). Default in all browsers. 92\%+ coverage. &
  Mature universal transport. Foundation for WebTransport and MoQ. \\
\midrule
WebTransport &
  Experimental only. Chrome disabled (M91--94). Firefox/Safari: none. &
  Chrome~97+, Firefox~114+. Safari: still absent. 82\% coverage. &
  Universal (Safari via Interop~2026). W3C Rec. Enables unbundled architecture and MoQ. \\
\midrule
WebCodecs &
  Chrome~94 only (Sept 2021). Firefox/Safari: none. First Public WD. &
  Chrome, Firefox~130+, Safari~26+. 94\% coverage. Still WD. &
  W3C Rec by 2027. AV1 HW encode ubiquitous. Custom media pipelines on all devices. \\
\midrule
MoQ &
  Did not exist. &
  IETF draft (v15, Oct 2025). No browser API. &
  RFC status. JS libraries on WebTransport + WebCodecs. Sub-second CDN streaming. \\
\bottomrule
\end{tabularx}
\caption{Browser capability status by standard across the full 2021--2031 timeline.}
\label{tab:summary}
\end{table}

\section{Conclusion}

This document has traced the real-time web communication stack across a full
decade, from its fragmented state in late 2021 through its current maturity in
March 2026 to its projected convergence by 2031.

In late 2021, the stack was transitional. WebRTC was the only finalized
standard with broad browser support, yet even it suffered from significant
cross-browser inconsistencies --- Safari's codec and feature limitations,
Chrome's ongoing SDP migration, and iOS reliability issues. QUIC had just been
standardized but was not yet universally enabled. HTTP/3 was shipping based on
draft specifications. WebTransport was effectively non-functional in any
browser. WebCodecs had just shipped in a single browser. MoQ did not yet exist
as a draft.

By March 2026, the picture is fundamentally different. The core protocol stack
--- from QUIC transport through HTTP/3 to WebRTC --- is standardized via
finalized RFCs and implemented by default across all major browsers. WebCodecs
provides cross-browser access to hardware-accelerated codecs at the frame
level. Two gaps remain: Safari's lack of WebTransport support and the
Chrome-only status of Encoded Transforms for end-to-end encryption.

The 2028--2029 projections close these gaps. WebTransport reaches Safari via
Apple's Interop~2026 commitment. Encoded Transforms become cross-browser.
WebCodecs achieves W3C Recommendation status. AV1 becomes a practical WebRTC
codec as hardware encoding expands. MoQ and RTP over QUIC reach RFC status,
establishing new standards-based paths for media transport. Hardware and
network improvements --- 40--50\% CPU gains since 2021, broadband speeds
approaching 150~Mbps --- ensure these standards land on infrastructure that
can fully exploit them.

By 2029--2031, the stack reaches full convergence. The ``unbundled''
architecture --- composing WebTransport, WebCodecs, and WebAssembly as
alternatives to monolithic \texttt{RTCPeerConnection} --- becomes practically
viable for specialized use cases like cloud gaming, scientific visualization,
and broadcast distribution. AV1 hardware encoding is ubiquitous across the
device spectrum. MoQ on WebTransport enables standards-based sub-second
streaming with CDN fan-out. Network bandwidth (200--300~Mbps fixed, 100+~Mbps
mobile) shifts the constraint from capacity to quality: 4K conferencing,
multi-stream layouts, and high-frame-rate screen sharing become routine.

The central observation is that standards maturation, hardware improvement, and
network growth are complementary, not independent. A standard that reaches
browser universality on hardware that cannot run it efficiently, or over
networks that cannot carry its output, delivers theoretical capability rather
than practical impact. The 2029--2031 timeframe is where all three curves
converge: every standard in this document will be finalized and universally
implemented, running on hardware with dedicated codec acceleration, over
networks with bandwidth to spare. This convergence is what distinguishes the
projected end state from a simple extrapolation of today's capabilities ---
and what makes the full real-time web communication stack, from transport to
codec to application, a resolved problem rather than a work in progress.

\newpage
\appendix
\renewcommand{\thesection}{\Alph{section}}
\renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\Alph{section}.\arabic{subsection}.\arabic{subsubsection}}
\input{appendix-info-theory}
\input{appendix-video-compression}

\end{document}
